---
title: "data607 project 1 msiracusa"
output: html_document
---

``` {r load data }
#load necessary packages
library(RCurl)
library(stringr)

#load txt file 
x <- 'https://raw.githubusercontent.com/miasiracusa/Data607/master/project1/tournamentinfo.txt'

#read into list
first_data <- read.delim(x, header=FALSE, stringsAsFactors=FALSE)

#look at data
head(first_data)

```
I want to work with the data in 2 sets, the first row for the competitors will populate the first set and the second row will populate the second set. I don't care about the headers, so I get rid of them.
```{r clean up}
#removed headers
second_data <- first_data[-c(1:3),]
head(second_data)
```

Now, I'll split the data into 2 sets and combine them back together using cbind.
```{r combine rows} 
mid1 <- second_data[seq(2, length(second_data), 3)]
mid2 <- second_data[seq(3, length(second_data), 3)]

full_data <- cbind(mid1, mid2)
head(full_data)
```
Since we are asked to focus on Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents, I only focus on those categories.
``` {r categorize data}
#use stringr package to extract the names
competitorname <- str_trim(str_extract(mid1, "([[:alpha:]]+\\s){2,3}"))

#state info is located in mid2, formated in two letter abbreviations
competitorstate <- str_trim(str_extract(mid2, "([[:alpha:]]+\\s){1}"))

#want only the number part of the pre rating, so extract the the numbers that are of length 3 or 4 and surounded by no digits. I first exctract all strings of digits, then extract the pattern i want
rating_pre <- as.numeric(str_extract(str_extract(mid2, "[^\\d]\\d{3,4}[^\\d]"), "\\d+"))

#want points in format "d.d"
totalpoints <- as.numeric(str_extract(mid1, "(\\d\\.\\d)"))

#want the numbers to end in |, so i extract all digits followed by |, then I extract all the digits. i keep getting an error "argument is not an atomic vector; coercing" but I don't know what that means or how to fix it (despite extensive googling)

opponent <- str_extract_all(str_extract_all(mid1, "\\d+\\|"), "\\d+")

#to calculate the average prerating of the competitors' opponents, i first make an empty column
avgpre_opponent <- c()

#then, I create a for loop that unlists (as numbers so i can find the mean) the preratings of each opponent for the respective players and finds the mean of this new list of numbers
  for (i in 1:length(mid1))
       {
    avgpre_opponent[i] <- round(mean(rating_pre[as.numeric(unlist(opponent[i]))]))
  }
```

I create the final data frame and create the csv file.
```{r final dataframe}
chess_data <- data.frame(competitorname, competitorstate, rating_pre, totalpoints, avgpre_opponent)

write.csv(chess_data, file = "chess_data.csv", row.names = FALSE)
```

