---
title: "data612 ass 2"
output: html_document
---

```{r setup, include=FALSE}
library("recommenderlab")
library("ggplot2")
library(tidyverse)
library(kableExtra)
library(knitr)
library(recommenderlab)
```

```{r load data}
movies <- read.csv("https://raw.githubusercontent.com/miasiracusa/Data607/master/Data612/Data%20612%20ass2/movies.csv")

ratings <- read.csv("https://raw.githubusercontent.com/miasiracusa/Data607/master/Data612/Data%20612%20ass2/ratings.csv")

#merge data frames by movieId

m <- merge(movies, ratings, by="movieId")
n <- data.matrix(m, rownames.force = NA)
data(MovieLense)
MovieLense
```



```{r plot, echo=FALSE}
hist(m$rating)
```
The data appears to be skewed to the left. It also appears that people prefer to rate movies with whole numbers. 

Next, I will only examine the most rated movies, so our data is more accurate.
```{r data clean}
#pick the most relevant data points (in this case the most rated movies)
movies <- MovieLense[rowCounts(MovieLense) > 50, colCounts(MovieLense) > 100]
```
Next, I will create a train dataset and a test dataset
```{r train test}
train <- sample(x = c(TRUE, FALSE), size = nrow(movies),
                      replace = TRUE, prob = c(0.8, 0.2))

movieTrain <- movies[train, ]
movieTest <- movies[!train, ]
```

Next, I will use the cosine method to test for similarity among the movies.
``` {r cosine}
similaritems <- similarity(MovieLense[, 1:25], method = "cosine", which = "items")
heatmap(as.matrix(similaritems))
```


The red indicates that there is a high similarity. This makes sense to appear along the diagonal, since the movies are being compared to themselves. The darker yellow indicates slgihtly more similar items.
Now we will create an item to item model using the IBCF method.We will start with the training set.
```{r rec system}
trainI2I <- Recommender(movieTrain, method = "IBCF", param=list(normalize = "Z-score", method="Jaccard")) 
```

```{r rec system B}
class(predicted)
predicted <- predict(object = trainI2I, newdata = movieTest, n = 10)
rec <- sapply(predicted@items, function(x) {colnames(movies)}) 
rec %>% kable() %>% kable_styling(full_width = T)
```

Above, we can see which movies are recommended to each user in each column.

For the second method, I will use
``` {r pearson}
similar_items <- similarity(MovieLense[, 1:25], method = "pearson", which = "items")
heatmap(as.matrix(similar_items))
```
This appears to return more similarity between items, as the yellows are all a lot darker.
Next I will use the UBCF method to create a recommender system I will start by creating the training set.
```{r rec system 2}
trainI2I2 <- Recommender(movieTrain, method = "UBCF", param=list(normalize = "Z-score", method="Jaccard")) 
```

```{r rec system 2B}
predicted <- predict(object = trainI2I2, newdata = movieTest, n = 10)
rec <- sapply(predicted@items, function(x) {colnames(movies)[x]}) 
rec %>% kable() %>% kable_styling(full_width = T)
```
From a quick glance, the second method appears to be much more efficient for this data.