---
title: "Data607 Assignment 3"
output: html_document
---

Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to problems 3 and 4 from chapter 8 of Automated Data Collection in R.  Problem 9 is extra credit.  You may work in a small group, but please submit separately with names of all group participants in your submission.

Here is the referenced code for the introductory example in #3:

#3a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

#First, I need to load the  stringr package, the data, and create the 'name' vector.
```{r setup}
library(stringr)

raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
     
  #separate the last names
    last_name <- str_extract(name,"[[:alpha:]]+($|,)")
    last_name <- str_extract(last_name, "[[:alpha:]]+")
    last_name
  #separate the first names
    first_name <- str_replace(name, "C.", "Charles")
    first_name <- str_extract(first_name, "([[:alpha:]]+( ))")
    first_name[is.na(first_name)] <- str_extract(name, "([[:punct:]])\\s[[:alpha:]]+")[is.na(first_name)]
    first_name <- str_extract(first_name, "[[:alpha:]]+")
 #make a data frame   
first_last <- data.frame(first_name, last_name)
first_last
```

 


#3b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
``` {r title}
str_detect(name, "[[:alpha:]]{2,3}[.]")
```
#3c) Construct a logical vector indicating whether a character has a second name.
``` {r 2nd name}

second_name <- str_replace(name, "[[:alpha:]]{2,3}[.]", " ")
second_name <- str_count(second_name, "\\w+")
second_name > 2

```

#4) Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

a) [0-9]+\\$
Strings of an undefined length consisting of digits and ending in a dollar sign.

``` {r example a}
a <- c("123456$")
unlist(str_extract_all(a, "[0-9]+\\$"))
```
b) \\b[a-z]{1,4}\\b

Strings of lowercase letters of length one to four and are word edges.
``` {r example b}
b <- c("egg", "table", "$weep")
unlist(str_extract_all(b, "\\b[a-z]{1,4}\\b"))
```

c) .*?\\.txt$

Strings that end in ".txt".
``` {r example c}
c <- c("egg.txt", "table", "$weep.txt")
unlist(str_extract_all(c, ".*?\\.txt$"))
```

d) \\d{2}/\\d{2}/\\d{4}
Strings that start with 2 digits, followed by a forward slash, followed by 2 digits, followed by a forward slash, followed by 4 digits.
``` {r example d}
d <- c("egg.txt", "04/19/1995", "$weep.txt", "02/28/1992")
unlist(str_extract_all(d, "\\d{2}/\\d{2}/\\d{4}"))
```
e) <(.+?)>.+?</\\1>
Strings that contain html formatting, or strings that contain "<", followed by anything "a", followed by ">", followed by anything, followed by "</", followed by "a", followed by ">".

``` {r example e}
e <- c("<b>egg.txt</b>", "<0>4/19/1995</0>", "<.txt>")
unlist(str_extract_all(e, "<(.+?)>.+?</\\1>"))
```


#9) The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others!
``` {r extra credit}
m <- c("<clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")
n <- unlist(str_extract_all(m, "[[A-Z].]{1,}"))
o <- paste(n, collapse = "")
p <- str_replace_all(o, "\\.", " ")
p
```


