---
title: "Data607 assignment 5 msiracusa"
output: html_document
---
First, I load the appropriate packages.
```{r load packages}
library(dplyr)
library(tidyverse)
library(tidyr)
```

Then, I load the data.
``````{r load data}
raw_data <- read.csv("https://raw.githubusercontent.com/miasiracusa/Data607/master/assignment5/data607%20week5.csv", na.strings = "")
raw_data
```

Then, I rename the columns and remove the empty row.
``` {r name data}
airport.data <- (raw_data)

names(airport.data)[c(1, 2, 3, 5, 6)] <- c("airline", "flight status", "Los Angeles", "San Diego", "San Francisco")

airport.data[5,1] <- c("ALASKA")
airport.data[2,1] <- c("AM WEST")

airport.data <- airport.data[-3,]

airport.data
```

Then, I transform the data using the tidyr and dplyr packages.
```{r transform data}

airport.data <- gather(airport.data, "airport", "flights", 3:7)
airport.data %>% mutate(flights = as.integer(flights))
airport.data <- spread(airport.data, "flight status", "flights")

```

Then, I perform an analysis using the dplyr package.
```{r analysis}

analysis <- airport.data %>%
  group_by(airline) %>%
        summarise(
            total.delayed = sum(delayed),
            total.ontime = sum(`on time`),
            total.flights = sum(delayed + `on time`),
            percent.delayed = (total.delayed / total.flights)
        )
analysis
```
We see that ALASKA airlines had proportionally more delayed flights than AM WEST.

